---
title: "Almond Yield Anomaly"
author: "Claudia Flores, Haley Grant, Shuhan Song"
date: "4/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE)
```

```{r}
## Read in packages
library(tidyverse)
library(janitor)
library(patchwork)
```

```{r}
## Read in data file 'clim.txt'
climate <- read_table2("clim.txt") %>% 
  clean_names() # Clean up column names in climate data frame
```

```{r}
## Source almond_model.R function into RMarkdown
source("./R/almond_model.R")
almond_model # Call the .R to view the function
```

```{r}
## Aggregating climate data

# Aggregate daily climate data frame into yearly and monthly averages for minimum and maximum temperature observations and sum for precipitation
climate_averages <- climate %>% 
  group_by(month,year) %>% # Grouping the same years and months observations within the year and month column to have results separated out into monthly and yearly results
  summarize(
  avg_tmax_c = mean(tmax_c), # Finding the average for maximum temperature
  avg_tmin_c = mean(tmin_c), # Finding the average for minimum temperature
  sum_precip = sum(precip) # Finding the sum for precipitation 
  )

```

```{r}
## Checking the model against three years and values: 2000 = 9.59, 2001 = 159.51, 2002 = 0.24

# Answer for 2000
answer_year_00 = almond_model(Tn_2 = 11.61473, P_1 = 57.318)
answer_year_00

# Answer for 2001
answer_year_01 = almond_model(Tn_2 = 8.822143, P_1 = 201.041)
answer_year_01

# Answer for 2002
answer_year_02 = almond_model(Tn_2 = 7.719107, P_1 = 20.338)
answer_year_02
```
```{r}
## Summarizing and visualizing the data

# Creating a data frame to plot second month minimum average temperatures and first month sum of precipitation for all years

climate_averages_1 <- climate_averages %>% 
  filter(month == 1) %>% # Filtering the first month for each year
  select(-avg_tmin_c) %>% # Removing the column named avg_tmin_c
  select(-avg_tmax_c) # Removing the column named avg_tmax_c

climate_averages_2 <- climate_averages %>% 
  filter(month == 2) %>% # Filtering the first month for each year
  select(year,avg_tmin_c,month) # Keep only two columns: year and avg_tmin_c

climate_averages_3 <- merge(climate_averages_1,climate_averages_2, by = "year") # Combine climate_averages_1 and climate_averages_2 data frames into one data frame called climate_averages_3

# Creating a plot of average minimum average temperatures for the second month of each year
min_temp <- ggplot(data = climate_averages_3, 
       aes(x = year , y = avg_tmin_c)) + 
  geom_point(color = "orange") + 
  labs(caption = "Figure 1: Average Minimum Daily Temperatures for February",
       x = "Year",
       y = "Temperature (Celcius)") +
  theme_light()
```

```{r}
# Creating a plot of sum of precipitation for the first month of each year
precipitation <- ggplot(data = climate_averages_3, 
       aes(x = year, y = sum_precip)) + 
  geom_point(color = "navy") + 
  labs(caption = "Figure 2: Annual Total Precipitation for January", 
       x = "Year",
       y = "Precipiation (mm)") +
  theme_light()
```


```{r}
## Creating a loop to insert variables from aggregated climate data into the almond_model function

# Create a data frame to store results from loop ran through almond_model.R function
results = as.data.frame(matrix
                          (nrow = 22, # Number of rows
                            ncol= 2) # Number of columns
                          )

# Loop
for (i in 1:nrow(climate_averages_3)) # For each value (i) in the rows in the climate_averages_3 data frame
  {
  results[i,2] = # Place the result for each value (i) in the results data frame in column 1
    almond_model( # Function
      Tn_2=climate_averages_3$avg_tmin_c[i], # Use values in the climate_averages_3 data frame column avg_tmin_c
      P_1=climate_averages_3$sum_precip[i] # Use values in the climate_averages_3 data frame column sum_precip
              )
  }

## Name your columns in your results data frame
names(results)[1] <- "year" # Name the first column year
names(results)[2] <- "almond_yield" # Name the second column almond_yield

# For your results data frame, add the correct year to the almond_yield result obtained from running the loop
results <- results %>% 
  mutate(year = c(1989:2010)) # Add in years starting from 1989 to 2010
```

```{r}
## Graphing results data frame that contain annual almond yield

# Graph
almond_yield_anomaly <- ggplot(data = results, 
                               aes(x = year, y = almond_yield)) + 
  geom_point(color = "brown") + 
  labs(caption = "Figure 3: Annual Almond Yield Anomaly (ton per acre)", 
       x = "Year",
       y = "Almond Yield Anomaly") +
  theme_light()
```


Almond yield anomaly increased its variability and frequency from 1988 to 2010 (**Figure 3**). In 1995, 1997, 2001, 2005, 2008, and 2010, almond yield anomalies were greater than 150 tons per acre. Comparing **Figure 3** to **Figure 1**, these years were years when the annual total precipitations for January were higher than the other years. This trend means that January precipitation has a strong and positive impact on almond yield anomaly. The highest anomaly was in 1995 when it reached `r round(max(results$almond_yield), 2)` tons per acre. Only in a few years were almond yield anomalies negative with the lower one being `r round(min(results$almond_yield), 2)` tone per acre in 1988, which has a much smaller magnitude than the positive anomalies. This gives a positive average almond yield anomaly of `r round(mean(results$almond_yield), 2)` tons per acre during 1988-2010. 

```{r}
(min_temp / precipitation) | almond_yield_anomaly
```

